### 2023-06-18 paper/reply_to_referee.txt
revised version

We thank the referee for the useful suggestions. We have addressed the points raised in the report, as explained below.

- "The authors have a novel definition of a lens, and a toy definition of lens finding. The results are likely very sensitive to these definitions (indeed their conclusion about source size biases is demonstrably at odds with other authors because they have different definitions of a detectable lens). The authors should investigate which of their results are robust to different definitions of a lens and a discovered lens, and which findings are not robust e.g. if they impose a minimum magnification cut, or only choose objects with a centroid within the caustic."

Reply - We added a subsection in the Discussion (6.6) to check how the key results vary with a magnification-based lens definition.

- "I also wonder how much source colour plays a role - we often think of lenses as red and lensed sources as blue. Given their source catalogue has colours, the authors might consider investigating if colour based finders can introduce a potential selection bias too."

Reply - A selection on colour is equivalent to a change in the properties of the underlying source population. Given the small differences in the strong lensing bias between galaxy-galaxy and galaxy-quasar lenses, which have different source redshift distributions, we expect the effect of colour selection to be similarly small.

- "This work doesn't seem to be a huge step forward compared to Collett & Cunnington 2016 (https://ui.adsabs.harvard.edu/abs/2016MNRAS.462.3255C/abstract). The authors don't seem to be aware of that paper and I think they need to rework their lensed quasar section substantially to compare their result to Collett & Cunnington, in a similar way to how they compare with Mandelbaum 2008 for galaxy-galaxy lenses."

Reply - We thank the referee for pointing out this important work that we had missed. Our paper, however, addresses a different question from the ones covered by Collett & Cunnington 2016: how galaxy-quasar lenses are different from galaxy-galaxy lenses. We hope that the revised text make this point more clear.

- "It would be good if this section had a more concrete application to time-delay cosmography. The main potential H0 bias in applying galaxy-galaxy results to lensed quasars is likely to be differences in the total density slope of the populations. A figure comparing these would be very useful. Ideally, the authors would also seek to quantify if the H0 shifts seen by Birrer 2020 (https://ui.adsabs.harvard.edu/abs/2020A%26A...643A.165B/abstract) are consistent with differences in the selection function of the SLACS and TDCOSMO lenses."

Reply - We are not convinced that comparing the total density slopes would help much with understanding the impact on H0: the slopes of time-delay lenses are typically measured with very high precision, so there are not many concrete cases in which having a prior on the density slope would help with the inference of H0. What currently dominates the uncertainty on H0 is the deviation from a pure power-law density profile, which Birrer et al. 2020 described by a mass-sheet transformation parameter. We added a subsection (5.2.2) to quantify the bias on lambda_mst, which can be directly translated into a bias on H0.

- "More context should be given to the size of the biases detected - are they likely to substantially change current or near-future conclusions about astrophysics/cosmology?"

Reply - We're a bit unsure about how to address this point. We believe that the quantitative statements in the conclusions (particularly bullet point 6 and 8) already give an answer to this question.

- "I found the introduction to get very technical, very quickly. Equation 1 isn't really an equation, more an illustration of what goes into the discovered lens population. I think a bit more qualitative text about lensing selection biases would be helpful before jumping in to equation 1."

Reply - We restructured the introduction accordingly.